Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming

### 2022-03-11 Paper/Manuscript_Stats&Comp/manuscript_STCO.tex
Edits to sec 4.8

\subsection{On computational cost in multi-dimensional cases}

The computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ in the approximation increases exponentially with $D$ (see eq.~(\ref{eq_m_multi})). The HSGP method can be computationally faster than the exact GP for larger datasets due the cubic scaling of the exact GP.

The HSGP model is definitely useful in $1$D and $2$D, even for highly wiggly functions, unless in $2$D the number of data points is very small ($n \lesssim 300$) where the regular formulation of a GP is already reasonably fast itself.

In the experiments of this paper for multivariate cases, mostly reported in Section~\ref{sec_computation_nD}, computation for HSGP model was faster than for the exact GP for most of the non-linear $2D$ functions or moderated-large sized ($n>1000$) $3$D datasets (even highly wiggly functions with large $m^* \approx 10000$); see Figure~\ref{fig_time_nD_log}).

@@

As it has been shown in case study III (Section~\ref{sec_caseIII}), the iterative diagnostic tool proposed in this study can be very useful for the multivariate cases as it allows to reduce $m^*$ to the minimum, reducing computational time drastically, and still getting an accurate approximation. Otherwise, user might be struggling in these multivariate cases.

When HSGP or exact GP is faster depend also on the implementation details that can have big effect on the constants, and thus any more detailed recommendations can only be made in the context of the specific software implementation.

>>>>

\subsection{The computational cost in the multi-dimensional setting}

The HSGP model is computationally superior $1$D and $2$D, even for highly wiggly functions, except when the number of data points so small ($n \lesssim 300$) that regular formulation of the GP is already reasonably fast itself. However, the computation time of the HSGP model increases rapidly with the number of input dimensions ($D$) since the number of multivariate basis functions $m^*=m_1\times \cdots \times m_D$ in the approximation increases exponentially with $D$ (see eq.~(\ref{eq_m_multi})). Yet, the HSGP method can still be computationally faster than the exact GP for larger datasets due the cubic scaling of the exact GP.

In our experiments with for multivariate problems (see Section~\ref{sec_computation_nD}), the computation time for the HSGP model was faster than for the exact GP for most of the non-linear $2D$ functions or moderated-large sized ($n>1000$) $3$D datasets (even highly wiggly functions with large $m^* \approx 10000$); see Figure~\ref{fig_time_nD_log}).

@@

As it has been shown in case study III (Section~\ref{sec_caseIII}), the proposed diagnostic tool can be very useful for multivariate problems as it allows one to reduce $m^*$ to the minimum, reducing computational time drastically, and still getting an accurate approximation.% Otherwise, the user might be struggling in these multivariate cases.

Whether HSGP or exact GP is faster will also depend on the specific implementation details, which can have big effect on the constants, and thus further detailed recommendations may depend on the specific software implementation.
